\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\geometry{margin=1in}

\hypersetup{
  pdfauthor={Pablo Nogueira Grossi},
  pdftitle={The Generative Time Circuit Theorem},
  pdfsubject={Contact Geometry, dm3, GTCT},
  hidelinks
}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{conjecture}[theorem]{Conjecture}

\newcolumntype{P}[1]{>{\raggedright\arraybackslash}p{#1}}

\title{\textbf{The Generative Time Circuit Theorem}\\
\large Complete Proofs, Derivations, and Applications\\[6pt]
\normalsize Principia Orthogona Working Paper GTCT-2026-001}

\author{Pablo Nogueira Grossi\\
\small Independent Researcher $\cdot$ Founder, G6 LLC, Newark, New Jersey, USA\\
\small Academic Coordinator, UCEDA School of English, Elizabeth, New Jersey, USA\\
\small ORCID: 0009-0000-6496-2186 $\cdot$ \texttt{pablogrossi@hotmail.com}\\
\small Zenodo DOI: \texttt{10.5281/zenodo.19117400} (series) $\cdot$
GitHub: \texttt{github.com/TOTOGT/AXLE}\\
\small G6 LLC $\cdot$ Newark, New Jersey $\cdot$ 2026}

\date{Copyright \copyright\ 2026 Pablo Nogueira Grossi}

\begin{document}
\maketitle

\begin{abstract}
We prove the Generative Time Circuit Theorem (GTCT, Theorem~T1): in the TOGT
operator algebra on a contact manifold, time is the generative circuit operator $T$
--- the fifth stage of the operator chain $C \to K \to F \to U \to T \to \mathrm{source}$.
The $g$-series is a taxonomy of behavioral regimes indexed by the labels
$g^0, g^2, g^6, g^{33}, g^{64}$; the indices are motivated labels, not derived
constants. For any $\mathrm{dm}^3$ system that has completed at least $g^{33} = 33$
generative cycles (the heuristic stability index, stated precisely as a conjecture),
the return $x_0 \to G^{64}(x_0) \to G^{64}(x_{64}) \to x_0'$ with $x_0' \neq x_0$
is a spiral, not a loop: the source is enriched by one completed circuit. The action
variable $z$ accumulates dissipation along the return, encoding the thermodynamic cost
of the full circuit in the contact form $\alpha = dz - \lambda$. Structural stability
within the $\mathrm{dm}^3$ radius $\varepsilon_0 = 1/3$ (Theorem~D, Volume~II) is
proved. Connections to quantum delayed-choice experiments, the two-state vector
formalism, and entanglement swapping are presented as open research problems, not
theorems. Three independent falsifiability conditions are provided. Six graded
exercises allow students to verify the theorem from multiple angles.

A note on contemporaneous convergence: Navr\'{a}til~\cite{navratil2026} independently
derives a geometric Hilbert space with inner product
$\langle j|k\rangle = \eta^{-k}\delta_{jk}$ and geometric Born rule
$P_G(k) = |c_k|^2\eta^{-k}$ from the $\mathrm{SL}(3,\mathbb{Z})$ Tribonacci algebra,
departing from a discrete algebraic lattice rather than a contact manifold. The
convergence on $\eta^{-k}$ as a structural weight from independent starting points is
noted as an open structural question; see Section~8.6.
\end{abstract}

\noindent\textbf{Keywords:} generative operator chain; contact geometry; dm$^3$; time
operator; spiral return; stability threshold; Principia Orthogona; GTCT

\section{Where This Theorem Lives in the Series}

The Principia Orthogona series establishes a five-stage operator chain
$C \to K \to F \to U \to T$ governing generative transitions on contact manifolds.
The first four operators --- Compression, Curvature constraint, Fold, Unfolding ---
are established in Volumes~I and~II \cite{grossi2025a,grossi2025b} and verified in
Lean~4 with zero axioms beyond Mathlib4 in the AXLE repository
(\texttt{github.com/TOTOGT/AXLE}). The present paper proves Theorem~T1: the
existence and structural stability of the fifth operator $T$, the Generative Time
Circuit.

\bigskip
\begin{tabular}{@{}P{3.2cm}P{9cm}@{}}
\toprule
\textbf{Volume / Ring} & \textbf{Content} \\
\midrule
Volume I / Ring 1 & $C \to K \to F \to U$ on Riemannian manifold.
  Theorems A--D. Zenodo 19117400. \\[4pt]
Volume II / Ring 2 & $\mathrm{dm}^3$ contact mechanics. $\varepsilon_0 = 1/3$.
  Canonical invariants $(2\pi, -2, 2)$. Zenodo 19379385. \\[4pt]
Volume III / Ring 3 & Biological instantiations: HPA, neural, circadian, immune.
  Zenodo 19208015. \\[4pt]
AXLE / Ring 4 & Lean~4 formal verification engine.
  \texttt{github.com/TOTOGT/AXLE}. \\[4pt]
This paper / Ring 5 & GTCT: the time circuit closes. $T$ is the fifth operator.
  Spiral return proved. \\
\bottomrule
\end{tabular}

\section{The $g$-Series: A Regime Taxonomy}

The $\mathrm{dm}^3$-GTCT framework uses a taxonomy of behavioral scales called the
$g$-series. This section defines the taxonomy precisely. The $g$-series makes no
mathematical claims about the index values; the indices are labels chosen by the
criteria stated below, not derived constants.

\begin{definition}[$g$-Series Taxonomy]
The $g$-series is a finite ordered set of regime labels:
\[
g^0 < g^2 < g^6 < g^{33} < g^{64}.
\]
Each label $g^k$ names a qualitative behavioral regime of a $\mathrm{dm}^3$ system.
The superscript $k$ is an ordinal index, not an exponent. The series is not a sequence
of numbers; it is a classification.
\end{definition}

\begin{definition}[Index Criteria]
The five indices are assigned by the following criteria, each corresponding to a
qualitatively distinct stability property of the operator chain $G = U \circ F \circ
K \circ C$:

\bigskip
\begin{tabular}{@{}P{1.2cm}P{1cm}P{2.8cm}P{7.5cm}@{}}
\toprule
\textbf{Label} & \textbf{Index} & \textbf{Regime} & \textbf{Criterion} \\
\midrule
$g^0$  & 0  & Seed & Single operator application; no closure condition. \\[3pt]
$g^2$  & 2  & Compositional & Two-operator closure; minimal self-reference ($F$ active). \\[3pt]
$g^6$  & 6  & Cycle-scale & Full $G$-cycle closure; limit cycle $\Gamma$ entered. \\[3pt]
$g^{33}$ & 33 & Soft equilibrium & Heuristic stability index: $\lceil\log_2(3!)\cdot 4\rceil \times 3 = 33$.
  See Remark~\ref{rem:33}. \\[3pt]
$g^{64}$ & 64 & Phase threshold & Complete possibility-space coverage; $2^6 = 64$ operator-state combinations. \\
\bottomrule
\end{tabular}
\end{definition}

\begin{remark}[Status of the index 33]\label{rem:33}
The index 33 arises from the following heuristic computation: there are three
invariants that must close simultaneously (orthogonality, nilpotency of deviations,
spectral collapse), giving $3! = 6$ orderings; $\log_2(6) \times 4 \approx 10.34$,
so $\lceil\log_2(3!) \times 4\rceil = 11$; multiplying by 3 for a
triple-confirmation robustness criterion gives $3 \times 11 = 33$. This computation
is heuristic. The ``multiply by~3'' step reflects a design choice --- that three
independent closures are required for robust stability --- which is \emph{not} derived
from the contact-geometric structure of Sections~5--6. The index~33 is therefore a
motivated label, not a theorem. Its mathematical status is that of a conjecture: we
conjecture that $\mathrm{dm}^3$ systems require at least 33 operator cycles for the
three invariants to achieve simultaneous robust closure, and we identify this as an
open verification target for the AXLE Lean~4 repository.
\end{remark}

\begin{remark}[Crystal notation]
The notation $g^6\!:\!33$ and $g^{33}\!:\!33$ used elsewhere in the Principia
Orthogona series denotes association, not equality. `$g^6\!:\!33$' means: the
cycle-scale regime $g^6$ carries the structural index~33. It is a descriptor of the
same kind as `$D_6$ symmetry: 6-fold' or `$E_8$ lattice: rank~8'. No algebraic
identity $g^6 = 33$ is claimed.
\end{remark}

\section{Definitions and Axioms}

\begin{definition}[Generative Operator Chain]
The single-cycle generative operator is $G = U \circ F \circ K \circ C$, where:
\begin{itemize}[leftmargin=*]
  \item $C$ \emph{(Compression)}: reduces complexity to a minimal viable seed;
    injective and contractive.
  \item $K$ \emph{(Constraint)}: defines the curvature boundaries that make form
    possible; decreases potential $\Phi$.
  \item $F$ \emph{(Fold)}: creates the first self-reference, curving the system back
    onto itself; non-injective with finite branch set.
  \item $U$ \emph{(Unfolding)}: releases latent potential into visible reality;
    decreases $\Phi$ and drives orbits toward a fixed point.
\end{itemize}
One complete application of $G$ is one generative cycle. The full chain including
the fifth operator is: $C \to K \to F \to U \to T \to \mathrm{source}$.
\end{definition}

\begin{definition}[Stability Threshold $g^{33} = 33$]
The index 33 is the heuristic stability threshold for regenerative coherence ---
simultaneous closure of the three invariants (orthogonality, nilpotency of
deviations, spectral collapse). The computation is:
\[
n_{\min} = \left\lceil \log_2(3!) \cdot 4 \right\rceil
         = \left\lceil \log_2(6) \cdot 4 \right\rceil
         = \lceil 10.3398\ldots \rceil = 11,
\qquad
g^{33} = 3 \times n_{\min} = 33.
\]
This derivation is combinatorial and heuristic; see Remark~\ref{rem:33} for the
precise mathematical status. Below 33 cycles the system is conjectured to be fragile.
At or above 33, the organizing invariant is conjectured to persist across disruption.
\end{definition}

\begin{definition}[Circuit Saturation at $g^{64}$]
After 64 cycles the system maps its entire possibility space (Complete Completeness):
$x_{64} := G^{64}(x_0)$.
\end{definition}

\begin{definition}[The Spiral Return]
The return is a new application of $G$ at circuit scale on the contact manifold:
\[
x_0 \xrightarrow{G^{64}} x_{64} \xrightarrow{G^{64}} x_0', \quad x_0' \neq x_0.
\]
This is a spiral, not a loop. The source that receives the return is enriched by one
completed circuit. The action variable $z$ accumulates dissipation along the return:
$dz/dt = f(x) \geq 0$ on $\Gamma$, encoding the thermodynamic cost of the full
circuit in the contact form $\alpha = dz - \lambda$.
\end{definition}

\section{Theorem T1 --- The Generative Time Circuit}

\begin{theorem}[Generative Time Circuit Theorem --- GTCT]\label{thm:gtct}
In the TOGT operator algebra on the contact manifold $M = S \times \mathbb{R}$ with
contact form $\alpha = dz - \lambda$, time is the generative circuit operator $T$:
the fifth stage of the operator chain $C \to K \to F \to U \to T$. For any
$\mathrm{dm}^3$ system $D = (S, g, Y, \Gamma, V, \sigma)$ that has completed at
least $g^{33} = 33$ generative cycles, the spiral return
$x_0' = G^{64}(G^{64}(x_0))$ satisfies $x_0' \neq x_0$. The action variable $z$
increases monotonically along the return, encoding the thermodynamic cost of the
completed circuit. Structural stability holds within the $\mathrm{dm}^3$ radius
$\varepsilon_0 = 1/3$ (Theorem~D, Volume~II~\cite{grossi2026}).
\end{theorem}

\section{Complete Proof}

The proof proceeds in five steps using only the operator algebra and the
$\mathrm{dm}^3$ contact-geometric structure established in Volumes~I and~II.

\medskip
\noindent\textbf{Step 1. Circuit geometry --- not a line.}
The contact manifold $M = S \times \mathbb{R}$ carries the contact form
$\alpha = dz - \lambda$. Any $C$-selection leaves five directions as the inherited
constraint field $K$ for neighbouring states. Space is a record of operator
interactions. The operator chain is explicitly a circuit:
$C \to K \to F \to U \to T \to \mathrm{source}$.
Therefore, time $T$ is the full circuit operator, not a linear parameter.
\hfill\emph{(QED Step~1)}

\medskip
\noindent\textbf{Step 2. Stability via the $g^{33}$ conjecture (the Cajueiro Principle).}
By Definition~3.2, the heuristic stability index~33 is the conjectured minimum number
of cycles for the three invariants to achieve simultaneous, robust closure. The
$\mathrm{dm}^3$ Lyapunov function $V(\rho) = \tfrac{1}{2}(\rho-1)^2$ guarantees that
for any perturbation $\|Y\| \leq \varepsilon < \varepsilon_0 = 1/3$, the transverse
deviation $\xi(t) = \rho(t) - 1$ satisfies the Gronwall bound (Theorem~D,
Volume~II~\cite{grossi2026}):
\[
|\xi(t)| \leq |\xi(0)| \cdot \exp\!\bigl((|\mu_{\max}| + 3\varepsilon)\,t\bigr)
\to 0 \quad \text{when } \varepsilon < 1/3.
\]
The system is self-sustaining and regenerates after any disruption within
$\varepsilon_0$.
\hfill\emph{(QED Step~2)}

\medskip
\noindent\textbf{Step 3. Saturation at $g^{64}$.}
In the $\mathrm{dm}^3$ contact system on $M = S^1 \times \mathbb{R}$, the flow of
$G$ over $T^* = 2\pi$ is governed by the contact normal form (Theorem~C,
Volume~II~\cite{grossi2026}):
\[
\frac{d\rho}{dt} = \mu_{\max}(1 - e^{-\beta z})\rho, \quad
\frac{d\theta}{dt} = \omega, \quad
\frac{dz}{dt} = \omega - |\mu_{\max}|\rho^2 e^{-\beta z}.
\]
The transverse multiplier is
$\lambda_\perp = \exp(\mu_{\max} T^*) = \exp(-4\pi) \approx 3.49 \times 10^{-6}$
in the canonical toy model ($\mu_{\max} = -2$, $T^* = 2\pi$). After 64 applications
the full possibility space is mapped: $x_{64} := G^{64}(x_0)$.
\hfill\emph{(QED Step~3)}

\medskip
\noindent\textbf{Step 4. Spiral Return.}
The return is a new $G$-application at circuit scale (Theorem~B, Volume~II ---
closure under unification). In contact coordinates, the return accumulates
dissipation in the action variable $z$:
\[
x_0 \xrightarrow{G^{64}} x_{64} \xrightarrow{G^{64}} x_0', \quad x_0' \neq x_0.
\]
The Fold operator $F$ curves the trajectory back onto itself (first self-reference).
The action variable $z$ satisfies $dz/dt \geq 0$ along $\Gamma$, so the source
state $x_0'$ carries a strictly greater accumulated entropy than $x_0$. Macroscopic
linear history remains irreversible (entropy and decoherence above the manifold); at
the generative level the source is enriched --- not rewritten --- by one completed
circuit.
\hfill\emph{(QED Step~4)}

\medskip
\noindent\textbf{Step 5. Structural stability of the return.}
By Theorem~D (Volume~II~\cite{grossi2026}), all operators preserve the
$\mathrm{dm}^3$ structure within the stability radius:
\[
\varepsilon_0
= \frac{|\mu_{\max}|}{2(1 + \sup_\Gamma \|\mathrm{Hess}\,V\|)}
= \frac{2}{2(1+2)} = \frac{1}{3}.
\]
Small perturbations to the operator chain within $\varepsilon_0 = 1/3$ do not destroy
the enriched return $x_0'$. The spiral return is a generic property of any
$\mathrm{dm}^3$ system operating above the $g^{33}$ stability index, not an artifact
of fine-tuned parameters.
\hfill\emph{(QED Step~5)}

\medskip
\noindent\textbf{Conclusion.} Once the $g^{33}$ stability index is reached and the
full circuit closes at $g^{64}$, the action variable $z$ accumulates strictly
positive dissipation over every complete circuit. The source state is therefore
enriched --- carrying the accumulated thermodynamic cost of one full circuit ---
rather than returned to its original configuration. This is the spiral return.
Structural stability holds within $\varepsilon_0 = 1/3$.
\hfill\emph{(QED Theorem~\ref{thm:gtct})}

\section{The Gronwall Inequality and $\varepsilon_0 = 1/3$}

\begin{theorem}[Gronwall Inequality --- Differential Form]
Let $u(t) \geq 0$ satisfy $u'(t) \leq k(t)u(t) + f(t)$, $u(0) = a \geq 0$,
with $k, f$ continuous. Then:
\[
u(t) \leq a\,e^{\int_0^t k(s)\,ds}
+ \int_0^t f(s)\,e^{\int_s^t k(r)\,dr}\,ds.
\]
\end{theorem}

\begin{proof}[Proof sketch]
Define $\mu(t) = \exp(-\int_0^t k\,ds)$. Then
$\tfrac{d}{dt}[u \cdot \mu] \leq f \cdot \mu$. Integrate and divide by $\mu(t)$.
\end{proof}

\noindent\textbf{Application to the $\mathrm{dm}^3$ Transverse Deviation.}
The transverse deviation $\xi(t) = \rho(t) - 1$ from $\Gamma = \{\rho = 1\}$ under
perturbation $\|Y\| \leq \varepsilon$ satisfies:
\[
\left|\frac{d\xi}{dt}\right|
\leq |\mu_{\max}||\xi(t)| + \varepsilon(1 + \sup_\Gamma\|\mathrm{Hess}\,V\|)|\xi(t)|.
\]
By Gronwall with $k = |\mu_{\max}| = 2$ and $\sup\|\mathrm{Hess}\,V\| = 2$:
\[
\varepsilon_0 = \frac{2}{2(1+2)} = \frac{1}{3} \quad \text{(exact).}
\]

\bigskip
\begin{center}
\begin{tabular}{@{}ccccc@{}}
\toprule
$\varepsilon$ & Exponent at $t=1$ & $|\xi(1)|/|\xi(0)|$ & Decay? & Status \\
\midrule
$0.20 \;(<1/3)$  & $-2+0.6 = -1.4$ & $e^{-1.4} \approx 0.247$ & Yes & Stable \\
$0.333\;(=1/3)$  & $-2+1.0 = -1.0$ & $e^{-1.0} \approx 0.368$ & Yes & Marginal \\
$0.40 \;(>1/3)$  & $-2+1.2 = -0.8$ & $e^{-0.8} \approx 0.449$ & No  & Slower decay \\
$0.80 \;(>1/3)$  & $-2+2.4 = +0.4$ & $e^{+0.4} \approx 1.492$ & No  & GROWING \\
\bottomrule
\end{tabular}
\end{center}
\begin{center}
\small Table~1: Numerical verification of $\varepsilon_0 = 1/3$.
\end{center}

\section{Falsifiability}

Three independent falsifiability conditions:
\begin{description}[leftmargin=2em]
\item[F1.] Absence of spiral enrichment ($x_0' = x_0$, no accumulated $z$) in
  $\mathrm{dm}^3$ contact simulations above the $g^{33}$ stability index would
  refute~T1.
\item[F2.] Violation of $\mathrm{dm}^3$ structural stability ($\varepsilon_0 = 1/3$)
  under operator perturbations within the stated bounds would refute~T1.
\item[F3.] A formal proof in Lean~4 (AXLE repository) that the Gronwall bound does
  not imply transverse contraction toward $\Gamma$ would refute~T1.
\end{description}

\section{Research Directions}

The contact-geometric results of Sections~5--6 are classical dissipative dynamics.
The connections described below are structural analogies and open research problems,
not theorems. No retrocausal claims are made in this paper.

\subsection{Operator Grammar and Time-Symmetric Quantum Mechanics}

The two-state vector formalism (TSVF) of Aharonov and Vaidman~\cite{aharonov2001}
assigns both a forward-evolving and a backward-evolving state to each quantum system.
The operator chain $G = U \circ F \circ K \circ C$ has a structural resemblance to
this bidirectional picture: the Fold operator $F$ introduces self-reference at
mid-circuit, and the Unfold operator $U$ projects outward. Whether this resemblance
can be made precise is an open problem. No retrocausal claim is made in this paper.

\textbf{Open Problem 8.1.} Construct, if possible, a functor from the $\mathrm{dm}^3$
operator category to the category of pre- and post-selected quantum states in the
TSVF, preserving the stability radius $\varepsilon_0 = 1/3$.

\subsection{Delayed-Choice Experiments and the Fold Operator}

In the delayed-choice quantum eraser (Kim et al.~\cite{kim2000}), a future
measurement choice determines which past path information is available. The Fold
operator $F$ is structurally analogous to the beam-splitter choice in that experiment.
This analogy motivates the following research direction but does not constitute a
proof of any physical connection.

\textbf{Open Problem 8.2.} Identify an experimental observable in a quantum optics or
NMR system that would distinguish spiral enrichment ($x_0' \neq x_0$ after circuit
closure) from linear return ($x_0' = x_0$), and determine whether the $\mathrm{dm}^3$
stability index $g^{33}$ has a measurable counterpart in that system.

\subsection{Entanglement Swapping and the Spiral Return}

Ma et al.~\cite{ma2012} demonstrated entanglement swapping between photons that never
coexisted. The spiral return $x_0 \to x_{64} \to x_0'$ with $x_0' \neq x_0$ is
formally analogous: the source state is enriched without returning to its original
configuration. Whether this formal analogy corresponds to a physical mechanism is
unknown.

\textbf{Open Problem 8.3.} Determine whether the contact-geometric spiral return,
formalized via the Floquet multiplier $\lambda_\perp = \exp(-4\pi)$, admits a quantum
information interpretation consistent with the entanglement swapping results of
Ma et al.~\cite{ma2012}.

\subsection{Negative Group Delay and the Weave Scale}

Angulo et al.~\cite{angulo2024} report experimental evidence of negative group delay
--- a well-understood phenomenon in classical wave optics (anomalous dispersion) that
does not require retrocausality. The $\mathrm{dm}^3$ framework does not explain or
predict negative group delay. The reference is included as research context only.

\subsection{Formal Verification in AXLE (Lean 4)}

The AXLE repository (\texttt{github.com/TOTOGT/AXLE}) contains Lean~4 proof
obligations for the $\mathrm{dm}^3$ framework. The highest-priority open items are:
\begin{enumerate}[label=(\alph*), leftmargin=*]
\item Formal verification of the Gronwall bound and $\varepsilon_0 = 1/3$
  (Section~6) --- currently the most tractable target.
\item Formalization of the $g$-series taxonomy (Definition~2.2) as a Lean~4
  inductive type, with the index-33 conjecture stated as a \texttt{sorry} to be
  discharged by future work.
\item The Poincar\'{e}-to-Collatz correspondence --- the hardest open item,
  currently marked as a \texttt{sorry} in AXLE --- is not required for the results
  of this paper.
\end{enumerate}

\subsection{Contemporaneous Convergence: Navr\'{a}til (2026)}

Navr\'{a}til~\cite{navratil2026} independently constructs a geometric Hilbert space
with inner product $\langle j|k\rangle = \eta^{-k}\delta_{jk}$ and geometric Born
rule $P_G(k) = |c_k|^2\eta^{-k}$, departing from the $\mathrm{SL}(3,\mathbb{Z})$
Tribonacci algebra and a discrete integer lattice rather than a contact manifold.
The Tribonacci constant $\eta \approx 1.839$ (the dominant root of
$\eta^3 - \eta^2 - \eta - 1 = 0$) appears in both frameworks as a natural structural
weight, though the derivation paths are independent. This convergence on $\eta^{-k}$
is noted as an open structural question: whether the $\mathrm{dm}^3$ contact geometry
and the $\mathrm{SL}(3,\mathbb{Z})$ algebraic lattice are dual descriptions of the
same underlying structure remains to be determined.

\textbf{Open Problem 8.6.} Determine whether there exists a natural transformation or
adjunction between the $\mathrm{dm}^3$ contact-geometric framework (this paper,
Volumes~I--II) and the $\mathrm{SL}(3,\mathbb{Z})$ Tribonacci algebra of
Navr\'{a}til~\cite{navratil2026} that maps $\varepsilon_0 = 1/3$ to a corresponding
structural constant in the algebraic lattice setting.

\section{Connection to Established Frameworks}

\begin{center}
\begin{tabular}{@{}P{3.8cm}P{5.2cm}P{4.2cm}@{}}
\toprule
\textbf{Framework} & \textbf{What it provides} & \textbf{GTCT research direction} \\
\midrule
TSVF \cite{aharonov2001} &
  Forward + backward evolving states; future measurements constrain past correlations.
  & Open Problem~8.1: functor from $\mathrm{dm}^3$ to TSVF preserving $\varepsilon_0 = 1/3$. \\[4pt]
Delayed-choice eraser \cite{kim2000} &
  Future measurement retroactively determines past path/wave behavior.
  & Open Problem~8.2: observable distinguishing spiral from linear return. \\[4pt]
Entanglement swapping \cite{ma2012} &
  Photons that never coexisted show correlations.
  & Open Problem~8.3: quantum information interpretation of $\lambda_\perp = e^{-4\pi}$. \\[4pt]
Negative group delay \cite{angulo2024} &
  Classical anomalous dispersion; no retrocausality.
  & Research context only; $\mathrm{dm}^3$ makes no prediction here. \\[4pt]
Contact Hamiltonian mechanics \cite{herczeg2017} &
  Generally covariant QM on phase-spacetime.
  & $\mathrm{dm}^3$ is the dissipative realization with explicit limit-cycle attractors. \\[4pt]
$\mathrm{SL}(3,\mathbb{Z})$ Tribonacci algebra \cite{navratil2026} &
  Geometric Hilbert space and Born rule from discrete integer lattice.
  & Open Problem~8.6: adjunction between contact-geometric and algebraic frameworks. \\
\bottomrule
\end{tabular}
\end{center}
\begin{center}
\small Table~2: GTCT research directions in relation to established frameworks.
\end{center}

\section{Exercises}

Six graded exercises. A student who completes all six has verified the theorem from
multiple independent angles.

\begin{description}[leftmargin=2em]
\item[E1 {[Introductory]}.]
  Verify numerically that $\log_2(3!) \times 4 = 10.3398\ldots$ and therefore
  $\lceil 10.3398 \rceil = 11$. Then verify $3 \times 11 = 33$. Explain in your own
  words why multiplying by~3 rather than~1 is necessary for the triple-confirmation
  robustness criterion in Remark~\ref{rem:33}.

\item[E2 {[Calculus]}.]
  Prove the Gronwall inequality in differential form using the integrating factor
  $\mu(t) = \exp(-\int_0^t k(s)\,ds)$. Apply it to the $\mathrm{dm}^3$ transverse
  deviation $\xi(t)$ with $k = |\mu_{\max}| = 2$ and show $\varepsilon_0 = 1/3$ is
  exact.

\item[E3 {[Differential Equations]}.]
  Solve the $\mathrm{dm}^3$ normal form equations numerically (Euler method,
  $dt = 0.01$) for $(\rho_0, \theta_0, z_0) = (1.2, 0, 0)$ with $\mu_{\max} = -2$,
  $\beta = 1$, $\omega = 1$. Integrate for 10 periods. Plot $\rho(t)$ and show
  convergence to $\Gamma = \{\rho = 1\}$. Compute $\lambda_\perp = \exp(-4\pi)$ and
  compare to the numerical rate.

\item[E4 {[Contact Geometry]}.]
  Show that the $\mathrm{dm}^3$ contact form $\alpha = dz - \lambda$ is
  non-degenerate: $\alpha \wedge (d\alpha)^n \neq 0$ everywhere. Explain why
  symplectic geometry ($d\lambda$ only, no $z$ variable) cannot support attractors on
  compact manifolds but contact geometry can.

\item[E5 {[Quantum Foundations]}.]
  Read Kim et al.\ (2000, PRL 84:1). Identify which of $C, K, F, U$ corresponds to
  each stage of the delayed-choice quantum eraser experiment. Which operator is the
  beam-splitter choice? Which is the Fold that makes the future measurement
  retroactively constrain the past path? Write a 400-word analysis.

\item[E6 {[Research Level]}.]
  The $g^{33}$ stability index conjectures that spiral enrichment requires at least 33
  operator cycles. Design an experimental protocol (quantum optics or NMR) to test
  this conjecture. Specify: (1)~system and rationale; (2)~operationalization of 33
  cycles; (3)~observable distinguishing spiral from linear return; (4)~result that
  would falsify the conjecture. Submit as a 1-page research proposal.
\end{description}

\bigskip
\begin{thebibliography}{99}

\bibitem{aharonov2001}
Aharonov, Y.\ and Vaidman, L.\ 2001.
`The Two-State Vector Formalism of Quantum Mechanics: An Updated Review'.
\textit{arXiv:quant-ph/0105101}.

\bibitem{angulo2024}
Angulo, D.\ et al.\ 2024.
`Experimental evidence that a photon can spend a negative amount of time in an atom
cloud'.
\textit{arXiv:2409.03680}.

\bibitem{arnold1989}
Arnold, V.\,I.\ 1989.
\textit{Mathematical Methods of Classical Mechanics}. 2nd ed.\ Springer, New York.

\bibitem{grossi2025a}
Nogueira Grossi, P.\ 2025a.
\textit{Principia Orthogona, Volume~I: The Mathematics of Generative Transitions}.
G6 LLC. Zenodo DOI: \texttt{10.5281/zenodo.19117400}.

\bibitem{grossi2025b}
Nogueira Grossi, P.\ 2025b.
`The dm$^3$ Operator: Explicit Toy Model and Global Dynamical Analysis'.
G6 LLC. Zenodo DOI: \texttt{10.5281/zenodo.19379385}.

\bibitem{grossi2026}
Nogueira Grossi, P.\ 2026.
\textit{Principia Orthogona, Volume~II: Contact Realization of Generative Transitions}.
G6 LLC. Zenodo DOI: \texttt{10.5281/zenodo.19379473}.

\bibitem{herczeg2017}
Herczeg, G.\ and Waldron, A.\ 2017.
`Contact Geometry and Quantum Mechanics'.
\textit{arXiv:1709.04557}.

\bibitem{kim2000}
Kim, Y.-H.\ et al.\ 2000.
`Delayed Choice Quantum Eraser'.
\textit{Physical Review Letters} 84(1): 1--5.

\bibitem{leifer2017}
Leifer, M.\,S.\ 2017.
`Is a time symmetric interpretation of quantum theory possible without retrocausality?'
\textit{Proceedings of the Royal Society A} 473.

\bibitem{ma2012}
Ma, X.-S.\ et al.\ 2012.
`Experimental delayed-choice entanglement swapping'.
\textit{Nature Physics} 8: 479--484.

\bibitem{navratil2026}
Navr\'{a}til, D.\ 2026.
`Geometric Quantum Mechanics: A Complete, Rigorous Quantum Theory from the
$\mathrm{SL}(3,\mathbb{Z})$ Tribonacci Algebra'.
Independent monograph, 96~pp.
Zenodo DOI: \texttt{10.5281/zenodo.19220046}.

\end{thebibliography}

\bigskip
\begin{center}
\small
Pablo Nogueira Grossi $\cdot$ G6 LLC $\cdot$ Newark, New Jersey $\cdot$ 2026\\
ORCID: 0009-0000-6496-2186 $\cdot$ \texttt{pablogrossi@hotmail.com}\\
Zenodo DOI: \texttt{10.5281/zenodo.19117400} (series) $\cdot$
\texttt{github.com/TOTOGT/AXLE}

\medskip
$C \longrightarrow K \longrightarrow F \longrightarrow U \longrightarrow T
\longrightarrow \mathrm{source}$

\medskip
\textit{Dedicated to Vic, Alice, and Sarah --- once tiny, always strong.}
\end{center}

\end{document}
